systemfit: A Package for Estimating\\ Systems of Simultaneous Equations in \proglang{R}

### 2007-06-06 paper/trunk/systemfit_estimation.tex
substituted "\COV" (defined in jss.cls) for all "\Cov"; removed definition(s) of "\Cov"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@710 591db2bd-e0a0-4d22-8c19-2192dbb7da31
substituted "\E [ . ]" (\E is defined in jss.cls) for all "E ( . )" and "E [ . ]"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@711 591db2bd-e0a0-4d22-8c19-2192dbb7da31

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{OLS} = \left( X'X \right)^{-1} X'y .
   \label{eq:ols}
\end{equation}
These estimates are efficient only if the disturbance terms are not
contemporaneously correlated, which means
$\sigma_{ij} = 0 \; \forall \; i \neq j$.
If the whole system is treated as one single equation,
the covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta}_{OLS} \right] = \sigma^2 \left( X'X \right)^{-1} ,
   \label{eq:olsCovSameSigma}
\end{equation}
where $\sigma^2 = E \left( u_{it}^2 \right)$.
This assumes that the disturbances of all equations have the
same variance.

If the disturbance terms of the individual equations
are allowed to have different variances,
the covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta}_{OLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} ,
   \label{eq:olsCovSingleSigma}
\end{equation}
where $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = \sigma_i^2 = E \left( u_{it}^2 \right)$.

@@

The Weighted Least Squares (WLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{WLS} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y ,
\end{equation}
where $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = \sigma_i^2 = E \left( u_{it}^2 \right)$.
Like the OLS estimates these estimates are only efficient
if the disturbance terms are not contemporaneously correlated.
The covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta}_{WLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
\end{equation}
If no cross-equation parameter restrictions are imposed,
the parameter estimates are equal to the OLS estimates.

@@

When the disturbances are contemporaneously correlated, a Generalized 
Least Squares (GLS) estimation leads to efficient parameter estimates.
In this case, the GLS estimator is generally called °Seemingly Unrelated Regression°
(SUR) estimator \citep{zellner62}.
It should be noted that while an unbiased OLS or WLS estimation requires only that 
the regressors and the disturbance terms of each single 
equation are uncorrelated
$( E \left[ u_i ' X_i \right] = 0 \; \forall \; i )$,
a consistent SUR estimation requires that all disturbance terms and all 
regressors are uncorrelated
$( E \left[ u_i ' X_j \right] = 0 \; \forall \; i, j )$.

The SUR estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{SUR} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y ,
   \label{eq:sur}
\end{equation}
where $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_{it} u_{jt} \right)$.
And the covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta}_{SUR} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
   \label{eq:surCov}
\end{equation}

@@

If the regressors of one or more equations are correlated 
with the disturbances ($E \left( u_i' X_i \right) \neq 0$),
OLS estimates are biased.
This can be circumvented by an instrumental variable (IV)
two-stage least squares (2SLS) estimation.
The instrumental variables for each equation $H_i$ 
can be either different or identical for all equations.
The instrumental variables of each equation may not be correlated with 
the disturbance terms of the corresponding equation 
($E \left( u_i' H_i \right) = 0$).

At the first stage new ('fitted') regressors are obtained by
\begin{equation}
   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X .
\end{equation}
At the second stage the unbiased two-stage least squares estimates
of $\beta$ are obtained by
\begin{equation}
   \widehat{\beta}_{2SLS} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
   \widehat{X}' y .
   \label{eq:beta2sls}
\end{equation}
If the whole system is treated as one single equation,
the covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta}_{2SLS} \right] = \sigma^2 \left( \widehat{X}'
   \widehat{X} \right)^{-1} ,
   \label{eq:2slsCovSameSigma}
\end{equation}
where $\sigma^2 = E \left( u_{it}^2 \right)$.
If the disturbance terms of the individual equations
are allowed to have different variances, 
the covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta}_{2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1} ,
   \label{eq:2slsCovSingleSigma}
\end{equation}
where $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = \sigma_i^2 = E \left( u_{it}^2 \right)$.

@@

The Weighted Two-Stage Least Squares (W2SLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{W2SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y ,
\end{equation}
where $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = \sigma_i^2 = E \left( u_{it}^2 \right)$.
The covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta}_{W2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1} .
\end{equation}

@@

If the regressors are correlated with the disturbances 
($E \left( u_i' X_i \right) \neq 0$) and
the disturbances are contemporaneously correlated, 
a Generalized Least Squares (GLS) version of the two-stage least squares
estimation leads to consistent and efficient estimates.
This estimation procedure is generally called °Three-stage Least
Squares° (3SLS) \citep{zellner62b}.

The standard 3SLS estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{3SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y ,
   \label{eq:3slsGls}
\end{equation}
where $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_{it} u_{jt} \right)$.
Its covariance matrix is
\begin{equation}
   \Cov \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1} .
   \label{eq:cov3sls}
\end{equation}
While an unbiased 2SLS or W2SLS estimation requires only that
the instrumental variables and the disturbance terms of each single 
equation are uncorrelated
$( E \left[ u_i' H_i \right]) = 0 \; \forall \; i )$,
\cite{schmidt90} points out that this estimator is only consistent 
if all disturbance terms and all instrumental variables are uncorrelated
$( E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$
with
\begin{equation}
   H =
   \left[ \begin{array}{cccc}
      H_1 & 0 & \cdots & 0\\
      0 & H_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & H_G
   \end{array}\right] .
\end{equation}
Since there might be occasions where this cannot be avoided,
\cite{schmidt90} analyses other approaches to obtain 3SLS estimators.
One of these approaches based on instrumental variable estimation
(3SLS-IV) is
\begin{equation}
   \widehat{\beta}_{3SLS-IV} = \left( \widehat{X}' \Omega^{-1} X
   \right)^{-1} \widehat{X}' \Omega^{-1} y .
   \label{eq:3slsIv}
\end{equation}
The covariance matrix of this 3SLS-IV estimator is
\begin{equation}
   \Cov \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \Omega^{-1}
   X \right)^{-1} .
\end{equation}
Another approach based on the Generalized Method of Moments (GMM)
estimator (3SLS-GMM) is
\begin{equation}
   \widehat{\beta}_{3SLS-GMM} = \left( X' H \left( H' \Omega H \right)^{-1}
   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y .
   \label{eq:3slsGmm}
\end{equation}
The covariance matrix of the 3SLS-GMM estimator is
\begin{equation}
   \Cov \left[ \widehat{\beta}_{3SLS-GMM} \right] =
   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1} .
\end{equation}
A fourth approach developed by \cite{schmidt90} himself is
\begin{equation}
   \widehat{\beta}_{3SLS-Schmidt} = \left( \widehat{X}' \Omega^{-1} \widehat{X}
   \right)^{-1} \widehat{X}' \Omega^{-1} 
   H \left( H' H \right)^{-1} H' y .
   \label{eq:3slsSchmidt}
\end{equation}
The covariance matrix of this estimator is
\begin{equation}
   \Cov \left[ \widehat{\beta}_{3SLS-Schmidt} \right] =
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega 
   H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X}
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} .
\end{equation}
The econometrics software EViews uses
\begin{equation}
   \widehat{\beta}_{3SLS-EViews} = \widehat{\beta}_{2SLS} + 
   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right) ,
   \label{eq:3slsEViews}
\end{equation}
where $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
as defined by (\ref{eq:beta2sls}).
EViews uses the standard 3SLS formula (\ref{eq:cov3sls}) to
calculate the covariance matrix of the 3SLS estimator.

>>>>

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{OLS} = \left( X'X \right)^{-1} X'y .
   \label{eq:ols}
\end{equation}
These estimates are efficient only if the disturbance terms are not
contemporaneously correlated, which means
$\sigma_{ij} = 0 \; \forall \; i \neq j$.
If the whole system is treated as one single equation,
the covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{OLS} \right] = \sigma^2 \left( X'X \right)^{-1} ,
   \label{eq:olsCovSameSigma}
\end{equation}
where $\sigma^2 = \E \left[ u_{it}^2 \right]$.
This assumes that the disturbances of all equations have the
same variance.

If the disturbance terms of the individual equations
are allowed to have different variances,
the covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{OLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} ,
   \label{eq:olsCovSingleSigma}
\end{equation}
where $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.

@@

The Weighted Least Squares (WLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{WLS} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y ,
\end{equation}
where $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.
Like the OLS estimates these estimates are only efficient
if the disturbance terms are not contemporaneously correlated.
The covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{WLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
\end{equation}
If no cross-equation parameter restrictions are imposed,
the parameter estimates are equal to the OLS estimates.

@@

When the disturbances are contemporaneously correlated, a Generalized 
Least Squares (GLS) estimation leads to efficient parameter estimates.
In this case, the GLS estimator is generally called °Seemingly Unrelated Regression°
(SUR) estimator \citep{zellner62}.
It should be noted that while an unbiased OLS or WLS estimation requires only that 
the regressors and the disturbance terms of each single 
equation are uncorrelated
$( \E \left[ u_i ' X_i \right] = 0 \; \forall \; i )$,
a consistent SUR estimation requires that all disturbance terms and all 
regressors are uncorrelated
$( \E \left[ u_i ' X_j \right] = 0 \; \forall \; i, j )$.

The SUR estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{SUR} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y ,
   \label{eq:sur}
\end{equation}
where $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = \E \left[ u_{it} u_{jt} \right]$.
And the covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{SUR} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
   \label{eq:surCov}
\end{equation}

@@

If the regressors of one or more equations are correlated 
with the disturbances ($\E \left[ u_i' X_i \right] \neq 0$),
OLS estimates are biased.
This can be circumvented by an instrumental variable (IV)
two-stage least squares (2SLS) estimation.
The instrumental variables for each equation $H_i$ 
can be either different or identical for all equations.
The instrumental variables of each equation may not be correlated with 
the disturbance terms of the corresponding equation 
($\E \left[ u_i' H_i \right] = 0$).

At the first stage new ('fitted') regressors are obtained by
\begin{equation}
   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X .
\end{equation}
At the second stage the unbiased two-stage least squares estimates
of $\beta$ are obtained by
\begin{equation}
   \widehat{\beta}_{2SLS} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
   \widehat{X}' y .
   \label{eq:beta2sls}
\end{equation}
If the whole system is treated as one single equation,
the covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{2SLS} \right] = \sigma^2 \left( \widehat{X}'
   \widehat{X} \right)^{-1} ,
   \label{eq:2slsCovSameSigma}
\end{equation}
where $\sigma^2 = \E \left[ u_{it}^2 \right]$.
If the disturbance terms of the individual equations
are allowed to have different variances, 
the covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1} ,
   \label{eq:2slsCovSingleSigma}
\end{equation}
where $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.

@@

The Weighted Two-Stage Least Squares (W2SLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{W2SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y ,
\end{equation}
where $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.
The covariance matrix of the estimated parameters is
\begin{equation}
   \COV \left[ \widehat{\beta}_{W2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1} .
\end{equation}

@@

If the regressors are correlated with the disturbances 
($\E \left[ u_i' X_i \right] \neq 0$) and
the disturbances are contemporaneously correlated, 
a Generalized Least Squares (GLS) version of the two-stage least squares
estimation leads to consistent and efficient estimates.
This estimation procedure is generally called °Three-stage Least
Squares° (3SLS) \citep{zellner62b}.

The standard 3SLS estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{3SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y ,
   \label{eq:3slsGls}
\end{equation}
where $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = \E \left[ u_{it} u_{jt} \right]$.
Its covariance matrix is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1} .
   \label{eq:cov3sls}
\end{equation}
While an unbiased 2SLS or W2SLS estimation requires only that
the instrumental variables and the disturbance terms of each single 
equation are uncorrelated
$( \E \left[ u_i' H_i \right]) = 0 \; \forall \; i )$,
\cite{schmidt90} points out that this estimator is only consistent 
if all disturbance terms and all instrumental variables are uncorrelated
$( \E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$
with
\begin{equation}
   H =
   \left[ \begin{array}{cccc}
      H_1 & 0 & \cdots & 0\\
      0 & H_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & H_G
   \end{array}\right] .
\end{equation}
Since there might be occasions where this cannot be avoided,
\cite{schmidt90} analyses other approaches to obtain 3SLS estimators.
One of these approaches based on instrumental variable estimation
(3SLS-IV) is
\begin{equation}
   \widehat{\beta}_{3SLS-IV} = \left( \widehat{X}' \Omega^{-1} X
   \right)^{-1} \widehat{X}' \Omega^{-1} y .
   \label{eq:3slsIv}
\end{equation}
The covariance matrix of this 3SLS-IV estimator is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \Omega^{-1}
   X \right)^{-1} .
\end{equation}
Another approach based on the Generalized Method of Moments (GMM)
estimator (3SLS-GMM) is
\begin{equation}
   \widehat{\beta}_{3SLS-GMM} = \left( X' H \left( H' \Omega H \right)^{-1}
   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y .
   \label{eq:3slsGmm}
\end{equation}
The covariance matrix of the 3SLS-GMM estimator is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS-GMM} \right] =
   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1} .
\end{equation}
A fourth approach developed by \cite{schmidt90} himself is
\begin{equation}
   \widehat{\beta}_{3SLS-Schmidt} = \left( \widehat{X}' \Omega^{-1} \widehat{X}
   \right)^{-1} \widehat{X}' \Omega^{-1} 
   H \left( H' H \right)^{-1} H' y .
   \label{eq:3slsSchmidt}
\end{equation}
The covariance matrix of this estimator is
\begin{equation}
   \COV \left[ \widehat{\beta}_{3SLS-Schmidt} \right] =
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega 
   H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X}
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} .
\end{equation}
The econometrics software EViews uses
\begin{equation}
   \widehat{\beta}_{3SLS-EViews} = \widehat{\beta}_{2SLS} + 
   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right) ,
   \label{eq:3slsEViews}
\end{equation}
where $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
as defined by (\ref{eq:beta2sls}).
EViews uses the standard 3SLS formula (\ref{eq:cov3sls}) to
calculate the covariance matrix of the 3SLS estimator.
